% PSM: Probabilistic Spatial Memory for Wearable Look-Back QA
% ECCV 2026 Wearables AI Workshop submission, v1 draft skeleton.
%
% This is a generic LaTeX skeleton, not the ECCV style itself —
% drop in eccv2026.sty / \documentclass{eccv2026} when the
% workshop's style file lands. The section content (sections 1-7,
% one .tex file each in this dir) is target-agnostic and will
% slot into the ECCV style directly with no rewrites.
%
% To compile locally (any TeX distribution):
%   cd journal/paper_drafts
%   pdflatex main.tex
%   bibtex main
%   pdflatex main.tex
%   pdflatex main.tex
%
% Section status (2026-06-04):
%   §1 Intro        v1 drafted
%   §2 Architecture v1 drafted
%   §3 Pipeline     v1 drafted
%   §4 Setup        folded into §5.1; standalone §4 omitted
%   §5 Results      v1 drafted (§5.5 multi-session locked;
%                                §5.8 Aria Gen 2 PENDING task #7)
%   §6 Limitations  v1 drafted
%   §7 Conclusion   placeholder; write last

% Springer LNCS template (ECCV's required format). When the
% workshop-specific eccv2026.sty drops, swap \documentclass{llncs}
% for \documentclass{eccv2026}; the body is otherwise compatible.
\documentclass[runningheads]{llncs}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage[hidelinks]{hyperref}
\usepackage{xspace}

% Macros used across sections.
\newcommand{\psm}{\textsc{PSM}\xspace}
\newcommand{\caponek}{\texttt{per\_cell\_cap}\xspace}

\title{Probabilistic Spatial Memory: A Bounded-Memory Retrieval
Substrate for Wearable Look-Back Question Answering}

% Author block intentionally blank for double-blind review.
\author{Anonymous}
\institute{Anonymous Institution\\\email{anon@example.com}}

\begin{document}

\maketitle

\begin{abstract}
A wearable AI assistant captures hours of egocentric video per day;
answering a look-back question (``where did I leave my keys?'')
requires retrieving the relevant moments under strict on-device
memory bounds. We propose \textbf{Probabilistic Spatial Memory}
(\psm), a streaming spatial-temporal index combining an H3 hex-grid
tessellation, a time-decayed ring buffer of HyperLogLog sketches per
cell, and reservoir-sampled CLIP-embedding exemplars. \psm{} exposes
a novel \caponek{} knob that smoothly trades strict spatial diversity
for embedding-similarity precision; the permissive endpoint
(\texttt{cap}=$K$) reduces to brute-force retrieval over all
in-reservoir exemplars, and on the highest-mobility session in our
Nymeria evaluation, \psm{} at the bounded-memory default paired with
a single MLLM rerank matches brute-force CLIP on localization
mIoU at one-fifth the candidate budget. The operating-point effect
generalizes across four sessions spanning street-, building-, and
room-scale mobility tiers. The C engine ($\sim$700 LoC) and Python
eval harness are released under MIT.

\keywords{Egocentric video \and Spatial memory \and Bounded memory \and
Look-back QA \and Retrieval substrate \and Wearable AI.}
\end{abstract}

% ---- Body sections -------------------------------------------------
\input{section_1_intro.tex}
\input{section_2_method.tex}
\input{section_3_pipeline.tex}
\input{section_4_related.tex}
\input{section_5_results.tex}
\input{section_6_limitations.tex}

% TODO: §7 Conclusion — write after §5 numbers stabilize.
\input{section_7_conclusion.tex}

\section*{Reproducibility}
\label{sec:reproducibility}
All experiments use the released PSM engine and Python evaluation
harness (MIT-licensed; URL withheld for double-blind review and
will be added in the camera-ready). Key reproduction commands:
\begin{itemize}\setlength{\itemsep}{0pt}\small
  \item \texttt{python scripts/sweep\_per\_cell\_cap.py} ---
    operating-point sweep on \texttt{shelby\_arroyo\_act0}
    (Tables~\ref{tab:headline}, \ref{tab:ex-ablation}).
  \item \texttt{python scripts/multisession\_per\_cell\_cap\_sweep.py}
    --- four-session generalization (Table~\ref{tab:multisession}).
  \item \texttt{python scripts/eval\_psm\_mllm.py ... --mllm gemini}
    --- PSM\,$\rightarrow$\,Gemini rerank
    (Table~\ref{tab:headline}, last two rows).
\end{itemize}

% ---- Bibliography -------------------------------------------------
\bibliographystyle{plain}
\bibliography{references}

\end{document}
