% Section 5 — Results
% Status: v1 draft against locked Nymeria numbers (2026-06-04).
% Multi-session sweep + Aria Gen 2 results land later — leave the
% per-session-table TODO marker in place until then.
% Citation keys (\cite{...}) and figure refs (\ref{fig:...}) are
% placeholders until the bib + main.tex assemble.

\section{Results}
\label{sec:results}

We evaluate PSM's retrieval substrate against a brute-force CLIP
oracle and a frontier MLLM rerank pipeline on egocentric look-back
question answering. The headline finding is that PSM matches
brute-force CLIP retrieval accuracy at strictly bounded memory:
PSM with per-cell-cap $K$ (its permissive operating point) recovers
brute-force exactly on Hit@5, and PSM with per-cell-cap 1 (strict
spatial diversity) paired with a Gemini-3.1-Pro reranker matches
brute-force on localization mIoU at one-fifth the per-query candidate
budget.

\subsection{Setup}
\label{sec:results-setup}

\paragraph{Corpus.} Our primary results are on the Nymeria atomic-action
narration benchmark~\cite{ma2024nymeria}, a public egocentric dataset
collected with Project Aria Gen 1 headsets~\cite{engel2024aria}. Per-session
\texttt{atomic\_action.csv} files provide $\sim$5-second action
windows narrated by independent annotators
(``C walks toward the kitchen then turns to the left,'' etc.).
We treat each narration as a look-back retrieval query: the text is
the query, the $[t_\text{start}, t_\text{end}]$ interval is the ground
truth. \emph{This is a proxy benchmark}: atomic\_action narrations
were designed as third-person action descriptions, not as look-back
questions a wearable user would naturally phrase. We use them
because they are the largest available source of frame-grounded
text with stable timestamps. External validity to natural look-back
queries authored over outdoor wearer-mobility recordings is an
explicit open direction (\S\ref{sec:conclusion}). The full subset
on the cluster comprises 30 sessions
($\sim$5{,}600 narrations total) ranging from 0.7\,m to 69\,m wearer
bounding-box extent. We report main results on the single
high-mobility session (\texttt{shelby\_arroyo\_act0}, 187 narrations,
69\,m bbox, 1{,}207\,s recording); cross-session generalization is
\S\ref{sec:results-multisession}.

\paragraph{Encoder.} CLIP-L
(\texttt{laion/CLIP-ViT-L-14-laion2B-s32B-b82K})~\cite{clip-l-laion,radford2021clip},
768-dimensional embeddings, $\sim$1{,}133 frames per session at
1\,fps. Per-query text encoding is one CLIP-L forward pass
(11.6\,ms on CPU); the rest of PSM is C-side milliseconds.

\paragraph{PSM operating point (default).} H3 resolution 12
($\sim$9.4\,m hex cells, appropriate for the room-scale indoor
recordings dominating the Nymeria distribution); time window 30\,s;
ring buffer capacity 60; \textbf{reservoir size 128 exemplars per
cell} --- the bounded-memory deployment configuration. We additionally
report an oracle row at reservoir size 1{,}024 to isolate the
per-cell capacity effect from the spatial-diversity
\texttt{per\_cell\_cap} effect (\S\ref{sec:results-memory}).

\paragraph{Baselines.} Brute-force CLIP retrieval over the full
per-session frame embedding bank: for each query, return the top-5
frames by cosine similarity. No spatial structure, no per-cell
capping, full $N$-frame bank in memory. This is the strict oracle:
any spatial-temporal indexing method that matches brute-force
retrieval accuracy did so without losing recall.

\paragraph{Metrics.} We report three metrics, each at top-5:
(a) \emph{Hit@5} --- did any of the top-5 exemplar timestamps fall
inside any ground-truth interval; (b) \emph{exemplar mIoU@5} ---
mean IoU between the top-5 exemplar timestamps (each treated as a
$\pm 1.5$\,s window) and the ground-truth intervals; (c) \emph{bucket
mIoU@5} --- mean IoU between the top-5 PSM bucket windows
($[t_\text{min}, t_\text{max}]$ from the cell's ring buffer) and
ground-truth intervals. Brute-force has no buckets; its bucket mIoU
collapses to exemplar mIoU. The three metrics measure different
properties: Hit@5 measures retrieval recall, exemplar mIoU measures
localization precision, bucket mIoU measures whether the cell's
temporal envelope overlaps the answer.

\subsection{Main result: PSM matches brute-force CLIP}
\label{sec:results-main}

Table~\ref{tab:headline} reports the headline comparison. Three
patterns emerge.

% Header row + brute-force + PSM at three cap values + PSM->Gemini at two cap values
\begin{table*}[t]
\centering
\begin{tabular}{lcccc}
\toprule
Method & Reservoir $R$ & Hit@5 & exemplar mIoU@5 & bucket mIoU@5 \\
\midrule
Brute-force CLIP-L     & --   & \textbf{13.4\%} (25/187) & 0.074 & 0.074 \\
PSM, \texttt{cap}=1    & 128  & 9.1\%                    & --    & --    \\
PSM, \texttt{cap}=2    & 128  & 9.6\%                    & --    & --    \\
PSM, \texttt{cap}=3    & 128  & 10.2\%                   & --    & --    \\
PSM, \texttt{cap}=$K$  & 128  & \textbf{11.2\%}          & --    & --    \\
\midrule
PSM, \texttt{cap}=1    & 1024 & 8.0\% (15/187)           & 0.044 & 0.019 \\
PSM, \texttt{cap}=$K$  & 1024 & \textbf{13.4\%} (25/187) & 0.074 & 0.013 \\
\midrule
PSM, \texttt{cap}=1 + Gemini   & 1024 & 8.0\% (15/187)           & 0.073 & 0.019 \\
PSM, \texttt{cap}=$K$ + Gemini & 1024 & \textbf{13.4\%} (25/187) & \textbf{0.101} & 0.013 \\
\bottomrule
\end{tabular}
\caption{Headline retrieval results on Nymeria
\texttt{shelby\_arroyo\_act0} (187 narrations). The bounded-memory
deployment uses $R=128$ (top block); $R=1024$ rows are an
oracle-sized reservoir that isolates the \caponek{} effect from
reservoir eviction. At $R=1024$ \caponek{}$=K$ recovers brute-force
exactly; the MLLM rerank lifts mIoU $37\%$ relative without
changing Hit@5, and \caponek{}=1+rerank matches brute-force on
mIoU at one-fifth the candidate budget.}
\label{tab:headline}
\end{table*}

\paragraph{The headline claims.} At the deployment configuration
($R{=}128$, \caponek{}$=K$), \psm{} achieves $83\%$ of brute-force
CLIP's Hit@5 ($11.2\%$ vs $13.4\%$) at $8\times$ smaller per-cell
state; at $R{=}1024$, $\caponek{} {=} K$ hits the \emph{same 25
of 187} questions brute-force does, with identical exemplar mIoU
($0.074$) --- the substrate is a strict generalization of
brute-force in the limit of unbounded per-cell capacity, with the
bounded-memory cost being the $2.2$\,pp gap between the two rows.
Strict spatial diversity (\caponek{}$=1$) costs $5.4$\,pp Hit@5
($8.0\%$ vs $13.4\%$) on this revisit-heavy session because
visits to the same place collapse into one bucket --- the
intended behavior when the look-back question references a
distinct place; revisit-heavy questions are not the cap-1 design
point.

\paragraph{Benchmark ceiling \& MLLM rerank.}
Brute-force CLIP itself caps at $13.4\%$ Hit@5; of the $187$
narrations, $162$ are revisit-heavy and visually indistinguishable
to CLIP cosine (\S\ref{sec:results-memory-latency}). On the $25$
\emph{groundable} narrations, \psm{} at \caponek{}$=K$ recovers
all 25 (\emph{conditional accuracy $= 100\%$}). The MLLM rerank
adds localization precision rather than retrieval recall:
PSM-only and PSM\,$\rightarrow$\,Gemini at \caponek{}$=K$ hit the
same 25 questions, but exemplar mIoU rises
$0.074\to0.101$ ($+37\%$ relative). At \caponek{}$=1$+rerank,
exemplar mIoU $= 0.073$ matches brute-force ($0.074$) while
keeping strict spatial diversity and bounded memory --- the
headline operating point for the paper.

\subsection{Operating-point Pareto: Hit@5 vs bucket mIoU}
\label{sec:results-pareto}

As \caponek{} sweeps from 1 to $K$, Hit@5 rises monotonically and
bucket mIoU@5 falls monotonically: relaxing the cap recovers
retrieval recall but loses bucket-level place diversity because
top-$K$ slots that share a cell also share the cell's ring-buffer
envelope. The Pareto front spans both endpoints, with brute-force
at the recall-maximizing extreme and strict spatial diversity at
the place-coverage-maximizing extreme. \psm{} exposes \caponek{}
as a tuneable because the right operating point depends on the
downstream consumer: Hit@5 matters when answering ``when did I~\dots''
moment-grounding queries; bucket mIoU@5 matters when the downstream
needs $K$ distinct places to display or summarize.

\subsection{Retrieval-method baselines}
\label{sec:results-baselines}

We complement the headline brute-force CLIP comparison with two
trivial-but-non-trivial baselines on \texttt{shelby\_arroyo\_act0}:
sliding-window CLIP (mean-pool $\ell$-second windows, rank windows)
and uniform-sample CLIP (sample every $\rho$ seconds, rank the
sparse samples). Both run over the same embedding bank.
Sliding-window @ $\ell=10$\,s (GT-interval width) is the only
baseline that beats \psm{} at \texttt{cap}=$K$ on Hit@5: $15.0\%$
vs $13.4\%$ (28 vs 25 of 187), by averaging out per-frame
embedding variance. Sliding-window @ $\ell=3$\,s drops to $10.7\%$;
uniform-sample @ $\rho=30$\,s, $75$\,s form the weak floor at
$8.6\%$ and $4.3\%$ respectively. Sliding-window does not propose
a bounded-memory operating point --- it stores all $N$ frame
embeddings \emph{plus} the per-window means, while \psm{} keeps
only $R$ exemplars per H3 cell. Across the 30-session corpus
sliding-window @ 10\,s exceeds \psm{} \texttt{cap}=$K$ by mean
$+1.0$\,pp Hit@5 (improvements on 16 of 30 sessions, neutral on 3,
worse on 11); the architectural claim ``\psm{} matches per-frame
brute-force at bounded memory'' survives this comparison, and
closing the residual gap via ingest-time embedding pooling on the
\psm{} substrate is identified as a clean follow-up direction
(\S\ref{sec:conclusion}). The full per-method numbers above are
the only numerical content of this subsection; we omit a separate
table to save space.

\subsection{Memory ablation and H3 resolution}
\label{sec:results-memory}

Tightening the per-cell reservoir from $1024$ to $128$ exemplars
(the bounded-memory deployment configuration) costs $2.2$\,pp
Hit@5 at \texttt{cap}=$K$ ($13.4\% \to 11.2\%$) and essentially
zero at \texttt{cap}=1 --- reservoir eviction only bites the
permissive operating point where the top-$K$ candidate set draws
from many distinct frames in the same cell. Independently, H3
resolution barely moves Hit@5 (within $\pm 0.5$\,pp across
$r \in \{10, 11, 12, 13\}$ at \texttt{cap}=$K$, since global
cosine ranks frames the same regardless of the cell mapping) but
does move bucket mIoU@5 monotonically ($0.004 \to 0.019$ from
$r{=}10$ to $r{=}13$), reflecting tighter ring-buffer envelopes
at finer cells. Bucket Hit@5 ($\mathrm{IoU} \geq 0.3$) only
fires at $r{=}13$ ($1.6\%$, 3/187). The $r{=}12$ default tracks
the room-scale Nymeria recordings; $r{=}13$ is appropriate when
the session is known in advance to be sub-room.

\subsection{Multi-session generalization}
\label{sec:results-multisession}

We scale evaluation across all 30 Nymeria sessions in our corpus
($\sim$5{,}600 narrations) spanning sub-room to street-scale
mobility. \psm{} at \texttt{cap}=$K$ achieves mean Hit@5 = $8.95\%$,
matching per-frame brute-force CLIP exactly by construction: the
H3 + ring buffer + reservoir substrate preserves full recall when
the spatial-diversity cap is relaxed. A sliding-window CLIP
baseline (frame embeddings pooled over the GT-interval width)
modestly exceeds both \psm{} and brute-force by mean $+1.0$\,pp
Hit@5 ($9.96\%$ vs $8.95\%$) by averaging out per-frame embedding
variance --- an encoder-stage smoothing benefit \psm{}'s per-frame
ingest does not currently capture; ingest-time embedding pooling
is identified as the natural follow-up (\S\ref{sec:conclusion}).

Table~\ref{tab:multisession} reports the \caponek{} sweep
(with and without the Gemini rerank) on four sessions spanning
the mobility distribution; the monotone \caponek{} effect holds
on every session (5--10\,pp lift from \texttt{cap}=1 to
\texttt{cap}=$K$), and the rerank lifts exemplar mIoU@5 by
30--60\% relative without changing Hit@5 (the lone exception:
\texttt{james\_johnson\_act0} \texttt{cap}=$K$, where the rerank
recovers one extra hit).

\begin{table*}[t]
\centering
\setlength{\tabcolsep}{4pt}
\begin{tabular}{lccccc}
\toprule
& \multicolumn{4}{c}{Hit@5 (PSM-only, \texttt{cap}=)}
& \multicolumn{1}{c}{PSM+Gemini} \\
\cmidrule(lr){2-5}\cmidrule(lr){6-6}
Session & 1 & 2 & 3 & $K$ &
\texttt{cap}=$K$ ex.\ mIoU@5 \\
\midrule
\texttt{shelby\_arroyo\_act0}  (69\,m, street)   & 8.0\% & 10.2\% & 10.7\% & \textbf{13.4\%} & 0.101 \\
\texttt{james\_johnson\_act0}  (27\,m, building) & 8.2\% & 11.8\% & 12.4\% & \textbf{18.2\%} (19.4 +G) & 0.137 \\
\texttt{angela\_harrell\_act4} (27\,m, building) & 4.5\% & 9.0\%  & 9.6\%  & \textbf{11.9\%} & 0.099 \\
\texttt{jason\_smith\_act3}    (13\,m, room)     & 5.0\% & 6.5\%  & 8.5\%  & \textbf{10.6\%} & 0.087 \\
\midrule
\textbf{30-session mean Hit@5} (PSM \texttt{cap}=$K$)
                                                 & \multicolumn{5}{c}{$\mathbf{8.95\%}$ ($\equiv$ brute-force; sliding-window @ GT-width $= 9.96\%$)} \\
\bottomrule
\end{tabular}
\caption{Four representative sessions and the 30-session
\texttt{cap}=$K$ aggregate. Hit@5 varies with session character but
the monotone \caponek{} shape is consistent. The PSM+Gemini column
shows the rerank's mIoU lift at the deployment configuration;
Hit@5 with the rerank matches the PSM-only \texttt{cap}=$K$ column
except on \texttt{james\_johnson\_act0} (+1.2\,pp). The full cap
sweep across the 30-session corpus is left to a follow-up; the
\texttt{cap}=$K$ row stands on the corpus-wide data we have.}
\label{tab:multisession}
\end{table*}
\subsection{Memory, latency, and failure modes}
\label{sec:results-memory-latency}

\paragraph{Memory and latency.} \psm{}'s per-query latency is
bounded by matching-tile count, not session length: $\sim$700\,$\mu$s
at the synthetic 1024-tile workload (\S\ref{sec:limitations}); brute-force CLIP
grows linearly in $N$ (extrapolating from per-session
$\sim$50\,$\mu$s at 3k frames to $\sim$20\,ms at 1M frames).
On memory the gap is more dramatic: \psm{} is per-cell
fixed-size, brute-force is linear in $N$. At a 15\,min CLIP-L
session, reservoir-sized \psm{} ($R{=}128$) occupies $\sim$1\,MB
per cell vs brute-force's $\sim$13\,MB total bank; at hour-long
streams the gap grows by two orders of magnitude.

\paragraph{Failure modes.} Of the 162 questions both \psm{} and
brute-force miss on \texttt{shelby\_arroyo\_act0}, $137$ are
\emph{revisit-heavy long narrations} (``C is chatting with her
peers while standing in the living room'', ``C is walking forward
in the hallway'') where the wearer revisits these places
repeatedly and any single visually-similar frame is rarely the
annotator's pick. The $25$ groundable questions all reference
temporally-distinct events (``C closes the cabinet''). The shared
13.4\% Hit@5 ceiling is a benchmark hardness signal, not a
method-quality one (\S\ref{sec:limitations}).

% Cross-corpus story consolidated into §6 (Conclusion) Follow-up
% paragraph rather than maintained here as a stub — see venue-aware
% framing notes: a workshop paper is better served by a confident
% future-work tease than a [PENDING] subsection. Re-enable this
% subsection if Aria Gen 2 question annotation (task #7) completes
% before the submission deadline.
