% Section 2 — PSM architecture
% Status: v1 draft. Describes the engine's three data structures
% (H3 cells / ring buffer of HLLs / reservoir-sampled exemplars)
% and the per_cell_cap selection rule. Anchored on the actual
% C-side surface in include/core/{spatial_memory,tile,ring_buffer}.h
% so any future reader can navigate from the prose to the code.

\section{Probabilistic Spatial Memory}
\label{sec:method}

PSM is a streaming spatial-temporal index. It ingests a stream of
$(t, \mathrm{lat}, \mathrm{lng}, \mathbf{e})$ tuples ---
timestamp, geographic position, and a $d$-dimensional CLIP
embedding --- and supports a bounded-memory top-$K$
similarity-search query at any time. The substrate is built from
three layered probabilistic data structures, each of which serves a
specific role in the bounded-memory property.

\subsection{H3 spatial tessellation}
\label{sec:method-h3}

We tessellate the wearer's geographic trajectory using
H3~\cite{uber-h3}, an open-source hexagonal hierarchical
geospatial indexing system. Every ingested $(t, \mathrm{lat},
\mathrm{lng}, \mathbf{e})$ tuple is mapped to its containing H3
cell at a chosen resolution $r$, and the per-cell state described
in the next two subsections is updated. The choice of $r$ governs
the spatial granularity:
H3 resolution 10 corresponds to $\sim$65\,m hex edges (street
scale); resolution 12 to $\sim$9\,m (room scale); resolution 13 to
$\sim$3.5\,m (sub-room). PSM exposes $r$ as a query-time parameter
because the right value depends on the wearer's mobility regime
(\S\ref{sec:results-multisession}).

The hexagonal tessellation has two properties relevant to PSM.
First, all six neighbors of a cell are equidistant from its
center, which makes k-ring neighborhood queries (``what was I
doing near here?'') trivial. Second, the cells are global ---
a single $(\mathrm{lat}, \mathrm{lng})$ position maps to the same
H3 cell across recordings, sessions, and devices --- which is the
key property for the federated-memory experiments we defer to
follow-up work.

\subsection{Time-decayed ring buffer of HyperLogLog sketches}
\label{sec:method-ring-buffer}

Each H3 cell maintains a ring buffer of $C$ HyperLogLog
sketches~\cite{flajolet-hll}. The buffer rotates on a configurable
time-window cadence (default 30\,s); ingesting a tuple writes the
hash of the frame's embedding into the current sketch and advances
the buffer pointer when the time-window elapses. Old sketches
expire as the ring wraps, implementing strict bounded-time
retention without an unbounded log.

The HLL sketches serve two purposes in PSM. First, they give a
cardinality estimate of distinct events the wearer experienced in
this cell over the last $C \cdot \mathrm{window}$ seconds, which
will be used by the counting question class
(``how many people did I see at the kitchen?'') in follow-up work.
Second, they give the cell's \emph{temporal envelope} ---
$(t_{\min}, t_{\max})$ across the non-expired sketches --- which
PSM returns as the bucket window for each top-$K$ candidate.
The bucket window is what the downstream MLLM consumes as its
``where in the recording should I look?'' hint.

\paragraph{Memory bound.} Each HLL sketch occupies
$2^p$ bytes at precision $p$ (default $p = 14 \Rightarrow$ 16\,KiB).
A cell's ring buffer is $C \cdot 16\,\mathrm{KiB} = 960\,\mathrm{KiB}$
at the default $C = 60$. The bound is per-cell and
time-window-bounded, independent of recording duration. We note
that for the v1 evaluation, only the buffer's $(t_{\min}, t_{\max})$
envelope is consumed downstream; the cardinality estimates are
forward-looking infrastructure for the counting question class
(\S\ref{sec:conclusion}). A pure $(t_{\min}, t_{\max})$ tracker
would drop per-cell state from 960\,KiB to a constant 16 bytes; we
defer that ablation to camera-ready since the counting use case
motivates carrying the HLL sketches as a v2 deliverable.

\subsection{Reservoir-sampled exemplars}
\label{sec:method-exemplars}

The HLL sketches give us counts and time envelopes but not the
embeddings themselves --- a top-$K$ similarity query needs to
score query-vs-frame cosines, which requires storing some of the
ingested embeddings. We use \emph{Algorithm R reservoir
sampling}~\cite{vitter-reservoir} per H3 cell, with a configurable
reservoir capacity $R$ (default 128). After ingesting $n > R$
frames into a cell, the reservoir holds a uniformly random
$R$-sized sample of all frames the wearer experienced there.

The reservoir is the bounded-memory substitute for the linearly-growing
embedding bank that brute-force CLIP keeps. At $R = 128$ and CLIP-L's
$d = 768$, each cell occupies $128 \cdot 768 \cdot 4 = 393\,216$
bytes $\approx 384$\,KiB --- a wearable that visits 100 distinct
cells in a
day stores $\sim$38\,MB total, vs $\sim$200\,MB for an embedding
bank over the same recording. The asymptotic gap grows with
recording length: brute-force is $\Theta(N)$; PSM is $\Theta(R \cdot
|\mathrm{cells}|)$ and $|\mathrm{cells}|$ is bounded by the
physical area the wearer covers, not the recording duration.

\paragraph{TurboQuant codec.} An optional exemplar-codec layer
(TurboQuant~\cite{turboquant}, our 2-/3-/4-bit implementation
in \texttt{src/core/exemplar\_codec.c}) compresses stored exemplars
5--10$\times$ at no Hit@5 cost compared to \texttt{raw} float32.
We report all main results with \texttt{raw} for clarity; the
codec is the natural next step for on-device deployment.

\subsection{Top-$K$ query with \texttt{per\_cell\_cap}}
\label{sec:method-query}

A top-$K$ similarity query takes a query vector $\mathbf{q}$
($d$-dimensional, matching the encoder), an optional
$(\mathrm{lat}, \mathrm{lng})$ center and a k-ring neighborhood
radius. The query semantics: ``return the top-$K$ candidates
$(t_i, \mathrm{cell}_i, \mathbf{e}_i)$ sorted by
$\cos(\mathbf{q}, \mathbf{e}_i)$ descending, with the cell's
ring-buffer-derived bucket window $(t_{\min}, t_{\max})_i$.''
Two query modes:
(a) \emph{global search} ($k$-ring $= -1$) iterates every
tile in the SpatialMemory; (b) \emph{local search}
($k$-ring $\geq 0$) iterates only the H3 cells in the $k$-ring
neighborhood around $(\mathrm{lat}, \mathrm{lng})$. The latter is
the right query mode for the ``where did I last see X near here?''
use case --- the PSM-only mode that needs zero embedding-space
text alignment.

\paragraph{Per-cell capacity bound.} The novel contribution at the
query layer is the \texttt{per\_cell\_cap} parameter. PSM's
top-$K$ aggregation step proceeds in two phases:
(i) for each H3 cell within the search radius, return the top-$\min$
exemplars within that cell, where $\min$ is $\min(\texttt{per\_cell\_cap},
\texttt{cell.reservoir.size})$;
(ii) sort the union of all per-cell results by similarity
descending and return the top $K$.
The selection structure is malloc-free at the per-cell layer
(insertion-sort into a fixed-cap buffer; bounded by
$\texttt{per\_cell\_cap}$, typically 1--$K$). The
\texttt{RingBuffer\_merge\_window} call --- the load-bearing
expensive op in the query path --- runs only on the top-$K$
\emph{surviving} cells, not on every scanned tile, which keeps
per-query latency sub-millisecond on session-scale workloads
(\S\ref{sec:results-memory-latency}).

The \texttt{per\_cell\_cap} parameter is the operating-point knob
that distinguishes PSM from brute-force CLIP retrieval.
\texttt{cap}=1 enforces strict spatial diversity: each top-$K$ slot
represents a distinct H3 cell. \texttt{cap}=$K$ removes the
per-cell constraint entirely, reducing to plain top-$K$ over the
union of reservoir exemplars across all in-radius cells.
The intermediate values trade off smoothly
(\S\ref{sec:results-pareto}). The default is \texttt{cap}=1
because, for the look-back QA use case PSM is designed for, the
question typically references a distinct place; the cap-$K$
permissive setting is the right tuning for revisit-heavy queries
where the question's answer is a moment, not a place.

\subsection{Streaming ingestion}
\label{sec:method-ingest}

PSM ingests frames one at a time. The ingest path per frame is:
(i) compute H3 cell from $(\mathrm{lat}, \mathrm{lng})$;
(ii) write the embedding hash into the current cell's
ring-buffer sketch and advance the buffer if the time-window
elapsed;
(iii) update the cell's reservoir via Algorithm R, which is
$O(1)$ per frame regardless of $n$.
No global re-indexing, no batch reorganization. The substrate
supports queries at any time after the first ingest.

The cluster-validated per-frame ingest cost is
$\sim$3\,$\mu$s amortized (Aria-internal benchmark, dominated by
HLL hash insertion). At 1\,fps Nymeria recordings the ingest
cost is negligible; at higher frame rates (60\,fps real-time
wearable streams) the per-frame budget is still well within
on-device limits.
