% Section 3 — Method: PSM as MLLM prefilter
% Status: v1 draft. Describes the three-stage pipeline (PSM search
% -> frame fetch -> MLLM rerank) and the frozen prompt template.
% Anchored on scripts/eval_psm_mllm.py and scripts/_mllm_client.py.

\section{Method: PSM as MLLM Prefilter}
\label{sec:pipeline}

The PSM substrate of \S\ref{sec:method} returns a top-$K$
candidate set for a query, but the candidates are
$(t, \mathrm{cell}, \mathrm{lat}, \mathrm{lng}, \mathbf{e})$
tuples --- not natural-language answers. For look-back QA
applications, we use PSM as a \emph{prefilter} that narrows the
search space, and a frontier MLLM as the \emph{reranker} that
picks the localized answer.

\subsection{Pipeline}
\label{sec:pipeline-stages}

The pipeline has three stages, each addressing a distinct failure
mode of the alternatives.

\paragraph{Stage 1: PSM top-$K$ retrieval.} Given the question
text $q$, embed it with the same encoder used for ingest
(CLIP-L for our experiments) and call
\texttt{SpatialMemory\_query\_similar} with the embedding as query
vector, the configured \texttt{per\_cell\_cap}, and $K = 5$. The
call returns 5 candidate $(t_i, t_{\min,i}, t_{\max,i},
\mathrm{cell}_i)$ tuples, sorted by cosine similarity descending.
This stage runs sub-millisecond per query
(\S\ref{sec:results-memory-latency}).

\paragraph{Stage 2: Frame fetch.} Each candidate $t_i$ identifies a
single frame in the original recording. We decode that frame
on-demand from the source video stream
(\texttt{aria01\_214-1.mp4} for Aria Gen 2, \texttt{data.vrs}'s
RGB stream for Nymeria) using either ffmpeg keyframe-seek (for
MP4 sources) or \texttt{projectaria-tools}'s indexed VRS reader
(for VRS sources). Frames are JPEG-encoded at quality 90 and
base64-embedded as \texttt{data:image/jpeg} URIs for the MLLM call.

The frame-fetch stage is the latency bottleneck of the eval
pipeline (1--5\,s/frame for VRS, 50--200\,ms/frame for MP4)
because it pays a video-seek cost per candidate. For a deployed
wearable assistant the stage would dissolve: the wearer's frames
are already in the device's frame buffer or in the local cache
that fed the PSM ingest loop, so the cost would be a memcpy
rather than a video decode. Our reported wall-clock numbers
should be read as eval-infrastructure overhead, not deployment
overhead.

\paragraph{Stage 3: MLLM rerank.} The 5 base64 frames plus the
question text are sent in a single API call to the frontier MLLM
(Gemini 3.1 Pro~\cite{google2026gemini} for the main results; the
same harness supports Claude 4.6 Opus by changing one flag, on the
same OpenAI-compatible proxy endpoint). The frozen prompt:

\begin{quote}\small\ttfamily
You are shown \{K\} candidate frames from a wearable camera
recording, numbered 1 to \{K\}. They are PSM-retrieved
candidates for the following look-back question:

  "\{question\}"

Pick the SINGLE frame number that best answers the question.
Respond with just the number (e.g.\ "3"). If none of the frames
clearly answer the question, respond with the number of the
frame you find most plausible --- do not refuse to choose.
\end{quote}

The prompt is the same across all sessions, all questions, both
MLLMs. We chose ``pick the index'' over ``write the answer''
because the eval metric is a temporal localization
(exemplar Hit@5 / mIoU), not free-text QA --- the MLLM's job is to
disambiguate among PSM's candidates, not to compose an answer.
The parsed integer reorders PSM's top-$K$ so the chosen frame is
in position 1 and the remaining four PSM candidates fill 2--$K$.

\subsection{Architectural decomposition}
\label{sec:pipeline-rationale}

The architectural claim of the paper is that PSM and the MLLM
contribute orthogonal properties to the retrieval task, and that
the composition matches a brute-force CLIP oracle on the metric
the MLLM downstream actually consumes.

\paragraph{PSM contributes \emph{recall}.} The substrate's job is
to return a candidate set that contains the answer. Empirically
(\S\ref{sec:results-main}), at \texttt{per\_cell\_cap}=$K$
PSM's top-5 contains the answer for the same 25/187 Nymeria
questions brute-force CLIP's top-5 does --- the substrate does not
lose retrieval signal vs the full embedding bank.

\paragraph{The MLLM contributes \emph{localization precision}.} The
reranker's job is to pick the most temporally-precise candidate
from PSM's top-5. Empirically (also \S\ref{sec:results-main}),
PSM\,$\rightarrow$\,Gemini at \texttt{cap}=$K$ achieves the same
Hit@5 as PSM-only at \texttt{cap}=$K$ (both 13.4\%) but lifts
exemplar mIoU@5 by 37\% relative (0.074\,$\rightarrow$\,0.101).
The MLLM is not recovering candidates PSM dropped; it is selecting
better-timed candidates from those PSM kept.

\paragraph{The composition.} At \texttt{cap}=1 + Gemini rerank,
PSM matches brute-force CLIP on the localization metric
(0.073 vs 0.074 exemplar mIoU@5) while keeping the cap-1
spatial-diversity guarantee. The bounded-memory substrate plus a
single MLLM call substitutes for the full embedding bank plus a
fine-grained top-$K$ search, on the metric a downstream MLLM
consumer cares about.

\subsection{Streaming considerations}
\label{sec:pipeline-streaming}

PSM ingest is $O(1)$ per frame
(\S\ref{sec:method-ingest}), so the substrate runs continuously
on-device alongside the wearer's normal recording loop. The
MLLM-rerank stage is the only one that requires a network call
to a remote API. We assume the wearer's assistant batches queries
to amortize the API cost; the eval-time per-query latency
($\sim$2--10\,s on Gemini, dominated by hidden thinking tokens for
reasoning models) is the right ballpark for a non-interactive
look-back use case (``end-of-day summary'', ``where did I leave
X?'') but would need a smaller / on-device MLLM for the interactive
real-time use case described in the workshop's CFP.

For the v1 results we report end-to-end wall-clock numbers from
the OpenAI-compatible proxy serving Gemini 3.1 Pro; we do not
claim PSM\,$\rightarrow$\,small-on-device-MLLM latency until
that substitution is measured. The workshop's Edge Computing
research area
(\S\ref{sec:intro} contributions list, third bullet) is the
natural target for that follow-up.
