% Section 1 — Introduction
% Status: v1 draft. Anchor on Wearables AI CFP topic-area alignment
% and the three-paragraph contribution structure:
%   (1) The problem: streaming egocentric retrieval is unbounded;
%       brute-force CLIP grows linearly in N.
%   (2) The architecture: PSM as a spatial-temporal index +
%       reservoir-sampled exemplars + ring-buffer time decay, with
%       per_cell_cap as the tuneable place-vs-similarity knob.
%   (3) The contribution: PSM matches brute-force at strict bounded
%       memory + spatial diversity, MLLM rerank composes with the
%       index multiplicatively (recall vs localization).

\section{Introduction}
\label{sec:intro}

A wearable AI assistant sees the world through the user's eyes,
continuously, for hours per day. To answer a look-back question
---``where did I leave my keys?'', ``what was that street sign?'',
``when did I last see Maria?''--- the assistant needs to find the
relevant moments from its own past. The straightforward solution is
to embed every frame the camera saw with a vision--language encoder
(CLIP, SigLIP, EVA) and run a top-$K$ cosine-similarity retrieval
over the resulting embedding bank. This works, but the embedding
bank grows linearly in recording duration: at a 1\,fps sampling and
CLIP-L's 768-dim float32 embeddings, an hour of egocentric video
occupies $\sim$10\,MB; a typical day, $\sim$200\,MB; the wearable's
on-device limit binds in days, not weeks. The substrate is not
appropriate for a bounded-memory streaming setting.

\paragraph{The architectural question.} What does the right
retrieval substrate look like for a wearable AI assistant? Three
properties matter. First, \emph{bounded memory}: per-cell state
that does not grow with recording duration. Second, \emph{spatial
locality}: the assistant should be able to answer
``what places did I visit?'' as cleanly as it answers ``what
moments matter?''; the index should reflect the user's actual
trajectory through physical space. Third, \emph{streaming
ingestion}: frames arrive one at a time, the index updates
in-place, and queries can be served at any time without
re-indexing.

\paragraph{Probabilistic Spatial Memory.}
We propose \textbf{PSM}, a streaming spatial-temporal index for
egocentric look-back retrieval. PSM combines three probabilistic
data structures over an H3 hex-grid tessellation of the wearer's
geographic trajectory:
(1) \emph{reservoir-sampled exemplars} per H3 cell preserve a
fixed-size sample of the CLIP embeddings the wearer captured in
that place;
(2) a \emph{time-decayed ring buffer of HyperLogLog sketches} per
cell tracks unique-event cardinality over recent time windows; and
(3) a \emph{per-cell capacity bound} (\texttt{per\_cell\_cap})
governs how many top-$K$ retrieval slots a single H3 cell may fill,
implementing strict spatial diversity at \texttt{cap}=1 and
permissive top-$K$-across-all-exemplars at \texttt{cap}=$K$.
The substrate is bounded in memory (fixed per cell, even as
recording duration grows unboundedly), streaming
(per-frame ingest, no global re-indexing), and trivially
queryable: a top-$K$ search returns $K$ candidate $(t, \mathrm{lat},
\mathrm{lng})$ triples in sub-millisecond C-side wall-time.

\paragraph{The PSM--MLLM prefilter pipeline.}
Modern multimodal LLMs (Gemini~3, Claude~4, GPT-4o) achieve strong
single-frame visual grounding but remain brittle on egocentric
video reasoning and temporal grounding over long-context
video~\cite{cheng2024videgothink,chen2025groundedmultihop,xu2026tgpo}.
For wearable look-back, this is exactly the failure mode: the
question asks ``when?'' and the MLLM cannot answer reliably without
seeing every frame, which the wearable cannot afford to ship to a
remote API. We propose PSM as a \emph{prefilter} for the MLLM:
PSM returns its top-$K$ spatial-temporal candidates over the
bounded-memory index; the MLLM reranks the $K$ candidate frames
and emits the localized answer. The architecture composes
multiplicatively: PSM provides the candidate set (recall), the
MLLM provides localization precision within the set, and the two
contributions are largely orthogonal.

\paragraph{Contributions.}
\begin{itemize}
  \item \textbf{Probabilistic Spatial Memory:} a streaming
    bounded-memory spatial-temporal index for egocentric video,
    combining H3 cells, time-decayed ring-buffer HLL sketches, and
    reservoir-sampled exemplars (\S\ref{sec:method}).
  \item \textbf{The \texttt{per\_cell\_cap} knob:} a new
    operating-point parameter that smoothly trades strict spatial
    diversity for embedding-similarity precision between two
    endpoints (\texttt{cap}=1 vs \texttt{cap}=$K$), reducing to
    plain top-$K$-over-all-exemplars at \texttt{cap}=$K$.
    Empirically, \texttt{cap}=$K$ recovers brute-force CLIP
    retrieval accuracy exactly on a 187-question Nymeria
    benchmark; \texttt{cap}=1 trades 5.4 percentage points of
    Hit@5 for spatial diversity (\S\ref{sec:results-main}).
  \item \textbf{PSM\,$\rightarrow$\,Gemini at \texttt{cap}=1 matches
    brute-force CLIP on localization mIoU on the highest-mobility
    session.} On \texttt{shelby\_arroyo\_act0} the bounded-memory
    operating point paired with a single Gemini-3.1-Pro rerank
    achieves exemplar mIoU@5 of 0.073, statistically indistinguishable
    from brute-force CLIP's 0.074. On lower-mobility sessions the
    permissive \texttt{cap}=$K$ operating point retains a margin
    (Table~\ref{tab:multisession}); the MLLM rerank lifts
    exemplar mIoU 30--60\% relative across all four sessions
    (\S\ref{sec:results-main}).
  \item \textbf{Multi-session generalization:} the
    \texttt{per\_cell\_cap} effect replicates across 4 Nymeria
    sessions spanning the street-, building-, and room-scale
    mobility tiers (\S\ref{sec:results-multisession}). The
    spatial-axis lift replicates further across
    \textbf{14 street-scale sessions} drawn from 3 distinct
    sensor stacks (SLOPER4D LiDAR + DJI Action2;
    Nymeria + LookOut~\cite{pan2025lookout} Project Aria MPS)
    and \textbf{3 encoder families} (laion CLIP ViT-L / ViT-bigG,
    Google SigLIP~2~\cite{tschannen2025siglip2}): 13/14 sessions
    pass a monotone-r10$\to$r12 acceptance with $\geq 4$\,pp
    Hit\,@\,5 lift on at least one encoder
    (\S\ref{sec:results-multi-corpus},
    Table~\ref{tab:multi-corpus-h3}). At matched candidate budget
    ($K=8$), \psm{} beats vanilla Gemini~3.1~Pro by $1.0\times$
    to $\infty$ on every one of these 14 sessions, median
    $\sim 4\times$ (\S\ref{sec:results-vanilla-mllm},
    Fig.~\ref{fig:f2-psm-vs-mllm}); four sessions land at
    vanilla MLLM Hit\,@\,5 $= 0\,\%$ because the uniform
    $K=8$ sample structurally never includes a frame inside any
    GT interval --- the failure mode \psm{}'s cell-targeted
    retrieval is designed to bypass.
  \item \textbf{Open-source release:} the C-side PSM engine
    ($\sim$700 LoC) and the Python extraction/eval pipeline are
    released under MIT (URL withheld for double-blind review;
    \S\ref{sec:conclusion} footnote).
\end{itemize}

The work touches three of the Wearables AI Workshop's areas:
Long-Context \& Real-Time Interactions (bounded per-query state),
Persistent Scene and Object Memory from Video (time-aware
spatial representation), and Efficient AI / Edge Computing (a
TurboQuant 2/3/4-bit exemplar codec compresses per-cell storage
5--10$\times$ at no Hit@5 cost, suitable for on-device deployment).
