% Section 4 — Related Work
% Status: v2 draft. Restructured per R3 venue notes — front-load
% the workshop audience's home turf (egocentric memory-augmented
% video LLMs), keep ANN/spatial-DB short. Citations curated against
% the workshop's CFP topic areas.

\section{Related Work}
\label{sec:related}

\paragraph{Egocentric video understanding and episodic memory.}
Ego4D~\cite{grauman2022ego4d} and its Natural Language Queries
benchmark established the look-back QA task on egocentric video at
3{,}000-hour scale; EgoSchema~\cite{mangalam2023egoschema}
contributed a long-form MCQ diagnostic that exposed the temporal-
grounding collapse motivating our PSM-as-prefilter pipeline. The
literal use case (``where did I leave my keys?'') was articulated
in earlier workshop work~\cite{barmann2022keys}; Di and
Xie~\cite{di2024grounded} introduced unified grounded-QA over long
egocentric videos with end-to-end neural grounding, and later
multi-hop variants make the same temporal localization pressure
explicit~\cite{chen2025groundedmultihop}. Recent egocentric
grounding work strengthens the case for a persistent spatial
substrate: Plizzari et al.~\cite{plizzari2025spatial} find that
spatial cognition is critical for locating objects across short and
long time scales, while fine-grained egocentric grounding benchmarks
such as EgoMask~\cite{liang2025egomask} expose the pixel- and
spatiotemporal-grounding challenges specific to first-person video.
Closest to our online setting, ESOM~\cite{manigrasso2026esom}
maintains a streaming object memory for episodic visual query
localization; its object-level memory is complementary to \psm{}'s
place-level H3 memory, and its reported 4\% online success rate
underscores the hardness of egocentric streaming retrieval. Our
contribution is structurally orthogonal: we provide a bounded-memory
spatial-temporal
substrate that any of these MLLM grounding methods can rerank, rather
than retraining a new grounding head. This PSM$\to$MLLM composition
is an instance of the video retrieval-augmented generation (VideoRAG)
pattern~\cite{jeong2025videorag}.

\paragraph{Memory-augmented video LLMs.}
MovieChat~\cite{song2024moviechat}, MA-LMM~\cite{he2024malmm},
Flash-VStream~\cite{zhang2024flashvstream}, and
LifelongMemory~\cite{wang2024lifelongmemory} all address bounded-
memory long-video understanding. More recent work tightens the bound:
HERMES~\cite{zhang2026hermes} arranges the KV cache itself as
hierarchical memory, InfiniPot-V~\cite{kim2025infinipotv} imposes a
training-free hard memory cap with 94\% reduction, and
SAVEMem~\cite{wu2025savemem} introduces a 3-tier semantic-aware
retention structure. These operate inside the MLLM's token space;
\psm{} operates upstream of it. They differ from \psm{} in two
important ways. First, they bound memory in \emph{token} space:
dense-to-sparse merging inside the LLM forward pass, KV-cache
compression, or recurrent state. \psm{} bounds memory in
\emph{geographic} space (H3 hex cells) and is therefore queryable
along a spatial axis that token-space memory cannot answer
(``what was I doing near here?''). Second, the token-space methods
require re-running the LLM on the compressed state; \psm{} runs
upstream of any MLLM, returning a top-$K$ candidate set that the
downstream model reranks without retraining. The two approaches are
complementary: \psm{} as front-end retrieval, token-space methods
as MLLM-internal compression. Emerging smart-glasses agent
frameworks such as Egocentric Co-Pilot~\cite{yang2026egocopilot}
combine temporal chain-of-thought with hierarchical context
compression for long-horizon first-person QA; \psm{}'s
bounded-memory index is the natural retrieval frontend such
agents query before invoking their MLLM reasoning stage.

\paragraph{Streaming search and spatial indexing.}
DiskANN~\cite{jayaram2019diskann} and SPANN~\cite{chen2021spann}
deliver billion-point ANN with bounded per-query cost over
near-static corpora; locally-adaptive vector
quantization~\cite{aguerrebere2024lvq} extends the streaming case
with per-vector compression. \psm{} differs in two ways: queries
optionally constrain to an H3 $k$-ring neighborhood (the
``near here'' spatial axis ANN does not expose), and ingestion is
fully streaming with bounded per-cell state that ages out via the
ring buffer. The H3 hex grid~\cite{brodsky2018h3} is a regular-grid
alternative to R-tree-style data-driven subdivisions~\cite{guttman1984rtrees};
its hexagonal geometry provides equidistant neighbors and a
globally-stable cell-id scheme, the latter being the basis for
the federated-memory extension we discuss in
\S\ref{sec:conclusion}. The probabilistic substrate
(HyperLogLog~\cite{flajolet2007hyperloglog}, reservoir
sampling~\cite{vitter1985reservoir}) is standard; the contribution
is in how the two compose with the H3 tessellation under the
\caponek{} query knob.

\paragraph{Vision-language encoders and the choice of substrate
input.} The retrieval quality of any embedding-bank approach,
including \psm{}, depends on the input encoder. We evaluate three
families that span the dominant pretraining recipes: laion's CLIP
ViT-L and ViT-bigG~\cite{cherti2023scaling} (softmax contrastive on
LAION-2B), and Google's SigLIP~2~\cite{tschannen2025siglip2}
(sigmoid pairwise on WebLI-100B, with explicit
localization-awareness training). Long-context CLIP
variants~\cite{zhang2024longclip}, denser-data
recipes~\cite{xu2023metaclip}, and ultra-scale variants such as
EVA-CLIP-18B~\cite{sun2024evaclip18b} are natural future swaps but
do not change the substrate. Dense visual encoders such as
DINOv2~\cite{oquab2024dinov2} and video-native encoders such as
V-JEPA~2~\cite{assran2025vjepa2} are complementary rerankers
when text--image similarity and visual salience disagree;
embedding-retrieval dimensionality limits~\cite{weller2025limit}
upper-bound any single-encoder substrate. We treat the encoder as
a swappable input (\S\ref{sec:pipeline}); the cross-encoder
generalization in \S\ref{sec:results-multi-corpus} shows the
spatial-axis claim is not a CLIP-family artifact.

\paragraph{Egocentric outdoor and street-scale corpora.}
Most egocentric benchmarks are indoor and short-trajectory; the
sensor stacks and outdoor capture needed for street-scale
spatial-memory evaluation are recent. Aria Gen~1 / Gen~2
glasses~\cite{engel2024aria,aria-gen2-pilot} ship with MPS SLAM
trajectories that we use as the geographic ground truth for the
Nymeria~\cite{ma2024nymeria} narrative-action sessions and for the
LookOut~\cite{pan2025lookout} pedestrian-navigation corpus
(downtown Burlingame / San Mateo / Foster City / Hillsdale /
Stanford walks). SLOPER4D~\cite{dai2023sloper4d} provides a
LiDAR-rigged head-mounted alternative with $\sim$300\,m urban
trajectories and frame-aligned RGB. Together these three corpora
cover the two dominant wearable-sensor stacks (Aria-style MPS,
helmet-LiDAR) and the street-scale mobility tier required to
exercise H3 resolutions 10--12; their licences
(CC~BY 4.0 / CC~BY-NC 4.0 / CC~BY-NC-SA 4.0 respectively)
permit research evaluation but restrict commercial use.
