% Section 6 — Limitations
% Status: v1 draft. Five limitations worth surfacing explicitly:
%   (1) Revisit-heavy long-narration ceiling
%   (2) Sub-room mobility exclusion (benchmark, not method)
%   (3) CLIP-77-token text truncation under-reports long-narration accuracy
%   (4) Single-corpus until task #7 lands
%   (5) Latency claim from synthetic workload, not Nymeria-scale validation

\section{Limitations}
\label{sec:limitations}

We surface five limitations of the v1 results so they are not
mistaken for stronger claims than they support.

\paragraph{Benchmark ceiling on revisit-heavy narrations.}
Both PSM and brute-force CLIP plateau at 13.4\% Hit@5 on the
\texttt{shelby\_arroyo\_act0} session (\S\ref{sec:results-main}),
and at similar single-digit values on the lower-mobility sessions
(\S\ref{sec:results-multisession}). The dominant failure mode is
\emph{revisit-heavy long narration}: the wearer enters the living
room many times during the recording, and each entry is
visually nearly identical, so the annotator's chosen ground-truth
interval is one of many indistinguishable matching frames. Both
methods converge on the visually-most-similar frame, which is
rarely the annotator's pick.
\emph{This is a benchmark hardness ceiling, not a method-quality
signal:} PSM and brute-force CLIP agree on the 25/187 questions
that \emph{are} groundable (\S\ref{sec:results-memory-latency}) and
disagree only on the inherently ambiguous remainder. Subsequent
benchmarks should either (a) include only narrations whose
ground-truth interval is the unique visual match in the recording,
or (b) score against \emph{any} acceptable revisit window rather
than the single one the annotator chose.

\paragraph{Sub-room mobility is excluded.}
Nymeria sessions with wearer bounding-box extent $<5$\,m
($\sim$13\% of our 30 candidate sessions; the
\texttt{nymeria\_slam\_displacement.py} probe gates this) are
excluded from the main result. On such sessions both \psm{} and
brute-force CLIP plateau near $1\%$ Hit@5 because the narrations
describe a space the wearer never left and the text--image
alignment cannot distinguish among visually indistinguishable
indoor frames. We treat sub-room mobility as a corpus
characteristic rather than a method limitation.

\paragraph{CLIP text encoder caps at 77 tokens.}
Nymeria \texttt{atomic\_action} narrations are full-sentence
descriptions (``C is walking forward in the hallway and turns left
toward the laundry closet, then opens the door with her right
hand''). Many tokenize to 80--120 tokens, exceeding the CLIP-L
text encoder's 77-token positional limit. We truncate at 77 tokens
at the tokenizer level; this drops trailing context that the
brute-force CLIP and \psm{} baselines both lose access to.
We tested whether a non-truncating text encoder would close this
gap, swapping in Long-CLIP-L~\cite{zhang2024longclip} (the same
ViT-L/14 architecture as CLIP-L, with positional embeddings
stretched from 77 to 248 tokens) and re-extracting both text and
image features on the headline session. The Long-CLIP swap
\emph{degraded} Hit@5 by $2.2$\,pp ($13.4\%\to11.2\%$), suggesting
that the truncated trailing context is auxiliary description that
adds noise rather than signal on this benchmark, and that
Long-CLIP's fine-tuning on long-caption COCO/Flickr-style pairs
shifts the embedding space away from what works for short-action
retrieval. Other long-context encoders (1024-token CLIP variants,
SigLIP~2~\cite{tschannen2025siglip2}) may behave differently; we
leave that broader sweep to future work.

\paragraph{Multi-corpus coverage.} The deepest single-session
analysis (Table~\ref{tab:headline}, the \caponek{} sweep,
the brute-force comparison) remains on Nymeria
\texttt{shelby\_arroyo\_act0}. The multi-corpus generalization
in \S\ref{sec:results-multi-corpus} (14 street-scale sessions
across SLOPER4D, Nymeria, LookOut $\times$ 3 encoder families)
addresses the obvious single-corpus / single-encoder objection
but uses Hit\,@\,5 as its sole metric --- we do not yet
report mIoU / counting / encoder-bypass numbers across the
full multi-corpus set. Per-session reservoir-size and
retention-horizon sweeps are also Nymeria-only at submission
time; replicating them across the 14-session set is a
straightforward follow-up that the engine and harness already
support.

\paragraph{One honest spatial-axis failure.}
Of the 14 street-scale sessions in
\S\ref{sec:results-multi-corpus}, \texttt{BurlingameDT4\_feb5}
(LookOut, 242\,m bbox) is the single session where \emph{all
three} encoders fail the monotone-r10$\to$r12, $\geq$4\,pp lift
acceptance. The trajectory is a short walk down a
parking-lot-flanked street where the captioned queries
(parked vehicles and street signage at small distances from
the wearer's path) collapse onto a handful of revisited cells
that no encoder discriminates above floor at the cell
granularity \psm{} operates on. This is the expected failure
mode of visually-homogeneous trajectories with
revisit-heavy query distributions; \psm{}'s spatial axis
is not magic and we report the negative case rather than
hide it.

\paragraph{Latency from synthetic workload.}
\psm{}'s reported $\sim$700\,$\mu$s per-query latency is from a
1024-tile dim-128 synthetic benchmark on H200; real Nymeria
sessions ($\sim$10 tiles at dim 1280, 128 exemplars) shift the
constants. The asymptotic frame-count independence claim holds
on both, but a real-session micro-benchmark is a follow-up.

\paragraph{Query-time pooling does not generalize.}
During method development, we explored a query-time reranking approach (\texttt{W=30} pooling) that averaged the embeddings of the $W$ frames surrounding each top-$K$ candidate before final selection. While this recovered performance on a small early-validation split, it failed to generalize across the 30-session corpus, behaving essentially as random noise relative to the per-frame baseline. We report this negative result to confirm that spatial-temporal smoothing must happen structurally at ingest time (via reservoir or ring buffer), not heuristically at query time.

\paragraph{What this paper does not claim.}
For clarity, we note five claims this work does \emph{not} make.
(a) We do not claim PSM is a generic MLLM grounding method;
PSM is a retrieval substrate, the MLLM does the grounding.
(b) We do not claim PSM solves the counting question class;
the time-decayed HLL sketches are diagnostic-only until the
real-counting scorer lands in a follow-up.
(c) We do not claim PSM is the only architecture for this
problem; we propose one shape, not the only one. (d) We do not
claim that per-frame brute-force CLIP is the strongest possible
CLIP-only baseline. Sliding-window CLIP at the GT-interval
width modestly exceeds both brute-force and \psm{} at \texttt{cap}=$K$
by mean $+1.0$\,pp Hit@5 across our 30-session corpus (with the
gap closing or inverting on 14 of those 30 sessions). This
encoder-stage smoothing benefit is real but small, and \psm{} does
not currently absorb it at its bounded-memory operating point ---
ingest-time embedding smoothing is the natural follow-up.
(e) We do not claim that CLIP-L is the strongest possible
encoder. Larger CLIP variants (bigG~\cite{cherti2023scaling},
EVA-CLIP-18B~\cite{sun2024evaclip18b}) and better-curated training
data (MetaCLIP~\cite{xu2023metaclip}) would shift absolute Hit@5
upward; localization-aware encoders such as
SigLIP~2~\cite{tschannen2025siglip2} are the natural next encoder
swap. Dense visual encoders such as DINOv2~\cite{oquab2024dinov2}
and video-native encoders such as V-JEPA~2~\cite{assran2025vjepa2}
are natural future re-rankers for cases where text--image
similarity and visual salience disagree, but the architectural
claim is encoder-agnostic.
