% Section 7 — Conclusion
% Status: v1 draft. Three paragraphs:
%   (a) restate the architectural finding
%   (b) honest scope: what generalized vs what didn't
%   (c) follow-up directions framed against the workshop's CFP

\section{Conclusion}
\label{sec:conclusion}

We presented \psm, a streaming bounded-memory spatial-temporal
retrieval substrate for egocentric look-back question answering.
The engine combines an H3 hex-grid tessellation, time-decayed
ring-buffer HyperLogLog sketches, and reservoir-sampled CLIP-
embedding exemplars, with a novel \caponek{} query knob that
smoothly trades strict spatial diversity for embedding-similarity
precision. The substrate's per-cell state is bounded independently
of recording duration; per-query latency is sub-millisecond after
the optimization in \S\ref{sec:method-query} amortizes the
per-tile RingBuffer merge over only the surviving top-$K$ rows.

\paragraph{What the evidence supports.}
On 30 Nymeria sessions, \psm{} at \caponek{}$=K$ matches per-frame
brute-force CLIP exactly by construction (mean Hit@5 $8.92\%$);
sliding-window CLIP at the GT-interval width modestly exceeds
both ($+1$\,pp). Paired with a Gemini-3.1-Pro rerank, strict
spatial diversity (\caponek{}$=1$) matches brute-force on
localization mIoU on the highest-mobility session
($0.073$ vs $0.074$) and lifts mIoU by 29--73\% relative across
the 4-session detail. Sub-room sessions co-fail both methods at
$\sim$1\% Hit@5 (benchmark ceiling, not method limitation).

The spatial-axis result generalizes across corpora and encoders.
On a 14-session street-scale set spanning three sensor stacks
(SLOPER4D LiDAR + DJI Action2, Nymeria + LookOut~\cite{pan2025lookout}
Project Aria MPS) and three encoder families (CLIP-L,
CLIP-bigG, SigLIP~2~\cite{tschannen2025siglip2}),
\textbf{13/14 sessions} pass a monotone-r10$\to$r12 acceptance
with $\geq 4$\,pp Hit\,@\,5 lift on at least one encoder.
SigLIP~2 is the strongest single encoder (11/14
PASS); two LookOut sessions are recovered by SigLIP alone where
both CLIP variants fail, and one session
(\texttt{BurlingameDT4\_feb5}) is an honest spatial-axis
failure on all three encoders (\S\ref{sec:limitations}). At
matched candidate budget ($K{=}8$),
\psm{} beats vanilla Gemini 3.1 Pro by $1.0\times$ to
$\infty$ on every one of these 14 sessions, median
$\sim 4\times$, with four sessions landing at vanilla MLLM
Hit\,@\,5 $= 0\,\%$ --- the structural failure mode that
uniform $K{=}8$ sampling has when GT intervals are
$\sim$3\,s wide but the uniform spacing is 12--80\,s.
The PSM$\to$MLLM rerank is permutation-invariant on Hit\,@\,5
and a bimodal Hit\,@\,1 lever: it tripled Hit\,@\,1 on
visually-rich \texttt{seq008\_running\_001} but hurt it on
action-narration Nymeria, so we deliberately do not promote
the rerank to a default.

The bounded-memory advantage is a deployment-scale property,
not a small-corpus one. \psm{}'s per-cell state is constant in
recording length; brute-force CLIP grows linearly in $N$.
The two methods cross at $N \approx 4400$ frames
($\sim$73\,min at 1\,fps, CLIP-L); past that the gap diverges
to $3\times$ at four hours and $6\times$ at eight hours of
recording.

\paragraph{Follow-up directions.}
Five extensions naturally fit the Wearables AI Workshop's
research areas. (a) \emph{Outdoor, GPS-grounded look-back QA}
on Aria Gen 2 Pilot's~\cite{aria-gen2-pilot}
\texttt{walk\_0}/\texttt{walk\_1} sessions ($\sim$300\,m wearer
trajectories with real GPS, H3 res 10 at street scale).
The engine and harness extend unchanged; the
operating-point claim becomes a cross-mobility-tier story
spanning $\sim$5\,m room scale to $\sim$300\,m street scale.
Question annotation is in progress at submission time; the
results land in the camera-ready or journal extension. This is
also the natural connection to the workshop's streaming
multi-turn Challenge 2 protocol. The recently released
Nymeria+~\cite{detone2026nymeriaplus}, which adds 3D bounding boxes
and object reconstructions, would further enable an object-level PSM
evaluation (the counting-unit ablation of E2).
(b) \emph{On-device MLLM rerank} closes
the loop on Edge Computing: replacing Gemini with a small on-device
vision-language model (e.g. Phi-3.5-Vision, SmolVLM) tests
whether the per-cell-cap + reranker composition holds when the
reranker itself is bounded by wearable compute. (c) \emph{Counting
via HLL cardinality}: the time-decayed ring-buffer sketches in
\S\ref{sec:method-ring-buffer} already track unique-event
cardinality per cell; surfacing them to the question answerer
extends \psm{} to a question class beyond moment-grounding
(``how many distinct people did I see at the kitchen today?'').
(d) \emph{Per-cell running-mean exemplars}: \S\ref{sec:results-multisession}
documents a $+1$\,pp Hit@5 gap between \psm{} and sliding-window
CLIP @ 10\,s that query-time and pre-ingest pooling did \emph{not}
close. Maintaining a single running-mean embedding per cell
alongside the reservoir (\(+\)one embedding per cell, \(\sim\)3\,KiB
at CLIP-L), and scoring both the top reservoir exemplar and the
cell mean at query time, is the smallest substrate change expected
to absorb the smoothing benefit without compromising
\texttt{exemplar\_t} precision; we defer the C-side implementation
and ablation to the camera-ready.
(e) \emph{Federated memory across wearers}: H3 cells are global,
so two devices visiting the same cell can merge their PSM state
via HLL union and reservoir sampling without sharing raw frames
--- a privacy-preserving foundation for the workshop's
``Persistent Scene and Object Memory'' theme at multi-user scale.

The engine ($\sim$700 LoC of portable C with HDF5 storage and
projectaria-tools/ffmpeg ingest) and the Python evaluation harness
are released under MIT.\footnote{Repository URL withheld for
double-blind review; added in the camera-ready. Key reproduction
entry points: \texttt{scripts/sweep\_per\_cell\_cap.py},
\texttt{scripts/multisession\_per\_cell\_cap\_sweep.py},
\texttt{scripts/eval\_psm\_mllm.py\,--mllm\,gemini}; the canonical
one-shot driver that regenerates the headline figures and Table~2
verdicts from the captured artifacts is
\texttt{scripts/reproduce\_paper.sh}.}
We hope \psm{} serves as a useful retrieval-substrate baseline
that future Wearable-AI assistants can deploy or improve on without
re-implementing the bounded-memory plumbing.
